---
title: "Homework 1"
author: "Preston Robertson"
date: "TBD"
output:
  
  word_document: default
  html_notebook: default
  pdf_document: default
---

## Regression Example

In this code, we demonstrate how to fit linear regression models using R. 

### Load data from library "datasets"
```{r message=FALSE}
library(datasets)
data(iris)
summary(iris)
head(iris)
```


### Visualize the data using pairwise scatter plots
```{r}
pairs(iris[1:4], main = "Iris Data", pch = 21, bg = c("red", "green3", "blue")[unclass(iris$Species)])
```

### 1. Fit a linear regression model using Sepal.Length as response and Sepal.Width as predictor
#### 1.B Fit the regression model
```{r}
lmfit.Sepal<-lm(formula=Sepal.Length ~ Sepal.Width, data=iris)
summary(lmfit.Sepal)
```


#### 1.C plot the data with the fitted regression line
```{r}
par(mfcol=c(1,1))
plot(iris$Sepal.Width, iris$Sepal.Length)
abline(lmfit.Sepal)
```
#### 2.A Based on your answer in 1c, fit a new model (Model 1) using Sepal.Length as response, Sepal.Width and Species as predictors.
```{r}
lmfit.Sepal<-lm(formula=Sepal.Length ~ Sepal.Width+Species, data=iris)
summary(lmfit.Sepal)
```
#### 3.A	Can you come up with another model that have better adj R2 than Model 1
```{r}
lmfit.Sepal<-lm(formula=Sepal.Length ~ Sepal.Width+Petal.Width+Petal.Length+Species, data=iris)
summary(lmfit.Sepal)
```

#### 3.3 Confidence interval of the parameters
```{r}
confint(lmfit.pedal,level=0.95)
```
#### 3.4 Residual plots
```{r}
par(mfcol=c(2,2))

# # cook's distance
# library(faraway)
# cook<-cooks.distance(lmfit.pedal) # check for influential points
# halfnorm(cook,3,ylab="Cooks dist")
# boxplot(cook)

# constant var
plot(fitted(lmfit.pedal),residuals(lmfit.pedal),xlab="Fitted",ylab="Residuals")
abline(h=0)

plot(fitted(lmfit.pedal),abs(residuals(lmfit.pedal)),xlab="Fitted",ylab="|Residuals|" )

# Normality
qqnorm(residuals(lmfit.pedal),ylab="Residuals")
qqline(residuals(lmfit.pedal))
hist(residuals(lmfit.pedal))

```

#### 3.5 Statistical tests
```{r message=FALSE}
library(lmtest)
library(lawstat)
library(nortest)

runs.test(residuals(lmfit.pedal))      # Runs test: test for randomness
dwtest(lmfit.pedal)                    # Durbin Watson test: test for independence (any autocorrelation?)
bptest(lmfit.pedal)                    # BP test: test for constant variance


ad.test(residuals(lmfit.pedal))         # Anderson-Darling test (normality)
shapiro.test(residuals(lmfit.pedal))    # Shapiro-Wilks test (normality)
```

