---
title: "iris_example3_classification"
author: "Wenmeng Tian"
date: "September 30, 2021"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
## classification Example

In this code, we demonstrate how to fit linear discriminant analysis and logistic regression models using R. 


### 1. Load data from library "datasets"
```{r message=FALSE}
library(datasets)
library(ggplot2)
data(iris)
summary(iris)
```

### 2. Visualize the data using pairwise scatter plots
```{r}
pairs(iris[1:4], main = "Iris Data", pch = 21, bg = c("red", "green3", "blue")[unclass(iris$Species)])
```

### 3. Fit a linear discriminant analysis model to predict iris species
#### 3.1 Split the data into training and testing set, and use training set to fit the LDA model
```{r}
library(MASS)
library(dplyr)
library(tidyverse)
library(caret)
library(e1071)
# Split the data into training and test set
set.seed(123)
training.samples <- iris$Species %>%
  createDataPartition(p = 0.8, list = FALSE)
train.data  <- iris[training.samples, ]
test.data <- iris[-training.samples, ]

# try to use functions in ggplot2 to visualize data distribution
density2 <- ggplot(data=iris, aes(x=Sepal.Width, fill=Species))
density2 + geom_density(stat="density", alpha=I(0.2)) +
  xlab("Sepal Width") +  ylab("Density") + ggtitle("Histogram & Density Curve of Sepal Width")

```
```{r}
# fit lda model
model.lda<- lda(Species~.,data=train.data)
print(model.lda)

```

#### 3.2 Use the testing set to evaluate the classification performance
```{r}
predictions.lda = data.frame(predict(model.lda, test.data))
confusionM.lda<-confusionMatrix(predictions.lda$class,test.data$Species)
#print(confusionM.lda$table)
print(confusionM.lda)
```

#### 3.3 Fit a QDA model and use testing set to evaluate classification performance
```{r}
model.qda<- qda(Species~.,data=train.data)
print(model.qda)
predictions.qda = data.frame(predict(model.qda, test.data))
confusionM.qda<-confusionMatrix(predictions.qda$class,test.data$Species)
print(confusionM.qda$table)
```
#### 3.4 Regularized discriminant analysis
```{r}
library(klaR)
cv_5_grid = trainControl(method = "cv", number = 5)
fit_rda_grid = train(Species ~ ., data = train.data, method = "rda", trControl = cv_5_grid)

fit_rda_grid
plot(fit_rda_grid)
```

### 4. Fit a logistic regression model to predict iris species
#### 4.1 Use the same training set in 3.1 to fit a logistic regression model
```{r}
library(nnet)
# Fit the model
model.logReg <- nnet::multinom(Species ~., data = train.data)
# Summarize the model
summary(model.logReg)
# Make predictions
predicted.logReg <- model.logReg %>% predict(test.data)
head(predicted.logReg)
# Model accuracy
mean(predicted.logReg == test.data$Species)

confusionM.logReg<-confusionMatrix(predicted.logReg, test.data$Species)
```
#### 4.2 Regularized logistic regression model

```{r}
library(glmnet)
library(fastDummies)
x<-train.data[,1:4] %>% as.matrix()
y<-train.data[,5] %>% as.matrix()

y_dummy<-dummy_columns(y)
colnames(y_dummy)[colnames(y_dummy) == "V1_setosa"] <- "setosa"
colnames(y_dummy)[colnames(y_dummy) == "V1_versicolor"] <- "versicolor"
colnames(y_dummy)[colnames(y_dummy) == "V1_virginica"] <- "virginica"


y_dummy1<-y_dummy[,2:4] %>% as.matrix()

# Perform 10-fold cross-validation to select lambda ---------------------------
lambdas_to_try <- 10^seq(-3, 3, length.out = 100)
# Setting alpha = 1 implements lasso regression
lasso_cv <- cv.glmnet(x, y_dummy1, family="multinomial", alpha =1, lambda = lambdas_to_try,
                      standardize = TRUE, nfolds = 10)
plot(lasso_cv)
lasso_model<-glmnet(x, y_dummy1, family="multinomial", alpha = 1,  lambda=lasso_cv$lambda.min, standardize = TRUE)
test_class_hat<-as.factor(predict(lasso_model, newx = as.matrix(test.data[,1:4]),type="class"))

print(confusionMatrix(test_class_hat,test.data$Species)$table)

```


